\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm} LFEUI - Investigation Proposal \\ \vspace{4mm} {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-1mm}

### 2023-12-02 text/proposta/prop.tex
vários

@@

\par In the following document we present a proposal for a scientific research being developed

@@

State clearly the aim of the proposed experiment together with the general and specific scientific background.

@@

Through the year there have been multiple techniques with the same purpose such as atomic spectroscopy and ?.
Particularly, the one we aim to use was introduced in ?.

@@

Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of
an experiment should be clarified with hosts before the proposal is submitted.

@@

The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with the samples
and the detectors.
The accelerator produces a beam of protons with a kinetic energy of up to 5 MeV which is deviated by the electromagnet
and hits the samples. 
The electromagnets deflect particles with different angles according to their charge-to-mass ratio.
This allows us to remove any unwanted particles from the beam and only let the protons pass through.
The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the
silicon.

In addition, their are connected to a electronic system that amplifies the signal and sends it to a computer where we
can see the results.
We aim to analyze the energy spectrum resulting from the reaction between the protons and the samples.
These spectra will unveil the presence of Li in the samples.

@@

\paragraph{Requirements}

@@

Specify the dates of the beamtime requested and the number of shifts. The beamtime requested should be in accordance with
the estimated duration of the experiment. Specify the place and the investigators who will guide the experiment.

\paragraph*{}

In order to complete our project we will need one full day of beamtime at the CTN laboratory. Alongside this day, we will
also meet with the investigators at CTN for another two mornings to discuss the project, before the experiment and the
results after the experiment:

@@

Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to
existing work on the topic.

@@

We expect to obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the
samples. 
For the glass samples, we expect to see a profile due to the particles that come from different depths of the sample and
therefore have lost different amounts of energy. In the implant samples, we expect to see a peak at the energy corresponding
to the energy of the alpha particles produced by the reaction.
In both cases, the samples is composed by other types of atoms such as silicon and boron, which will also emit charged
particles.

Other phenomena such as Rutherford Backscattering and Elastic Scattering are also expected to occur.

@@

Justify the beam time you request and give technical reasons which make the apparatus necessary for your experiment.
Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential
details.

>>>>

\setlist[enumerate]{itemsep=3pt}

@@

\par In the following document we present a proposal for a scientific research being developed at CTN Lisbon. We will...

@@

@@

This detection method is based on the nuclear reaction between the element we want to detect ($\ce{^7Li}$) and a proton, as follows:

\begin{equation}
  \ce{^7Li + ^1H -> ^8Be -> 2\alpha}
  \label{eq:reaction}
\end{equation}

The reaction results in the production of a $\ce{^8Be}$ nucleus which then rapidly decays into two alpha particles. These are the particles we detect.

According to the incident proton's energy and the scattering angle of the detectors ($\theta = 165 $ degrees) we were able to use the NRA Calculator \cite{NRAEnergyCalc} to calculate the energy of the alpha particles resulting from the reaction as well as it's $Q$ value:

\begin{equation}
\begin{split}
  E_{\alpha_1} &= 7.65320 \text{ MeV} \\
  E_{\alpha_2} &= 10.89305 \text{ MeV} \\
  Q &= 17.34625 \text{ MeV}
\end{split}
\label{eq:energies}
\end{equation}

The formula this website uses to compute the energies is the one on the book \cite{krane1988}.

@@

@@

The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with both the samples and the detectors.
The accelerator produces a beam of protons with a kinetic energy of up to 5 MeV which is deviated by the electromagnet and hits the sample. 
The electromagnets deflect particles with different angles according to their charge-to-mass ratio and by that allow us to remove any unwanted particles from the beam and only let the protons pass through.
The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the silicon.

In addition, their are connected to a electronic system that amplifies the signal and sends it to a computer where we
can see the results.

The beam line is under a vacuum of around $10^{-7}$ mbar in order to avoid the protons colliding with other particles and losing energy. 
They reach the sample with the specified kinetic energy and cause a nuclear reaction whose products are the charged particles detected.
We aim to analyze the energy spectrum of these particles.

The following diagram shows the experimental set-up:

\begin{center}
  \includegraphics[width=0.95\linewidth]{../images/scheme.jpeg}
  \captionof{figure}{Experimental set-up.}
  \label{fig:setup}
\end{center}

@@

@@

In order to complete our project we will need one full day of beamtime at the CTN laboratory \cite{IST}. Alongside this day, we will
also meet with the investigators at CTN for another two mornings to discuss the project, before the experiment and the
results after the experiment:

@@

We will obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the
samples. 
For the glass samples, we expect to see a profile due to the particles that come from different depths of the sample and
therefore have lost different amounts of energy. In the implant samples, we expect to see a peak at the energy corresponding
to the energy of the alpha particles produced by the reaction.
In both cases, the samples is composed by other types of atoms such as silicon and boron, which will also emit charged
particles.

\paragraph*{Energy Values} The alpha particles are supposed to have the energies calculated in equation \ref{eq:energies} ($E_{\alpha_1}$ and $E_{\alpha_2}$). 
In the glass sample, we expect to see an energy profile before each peak and in the implant sample we only expect to see the peaks.

@@

Other phenomena such as Rutherford Backscattering and Elastic Scattering are also expected to occur. These are responsible for energy peaks in both sample's spectra at lower than the expected energy values.

@@

By identifying the presence of light isotopes at low energies ($<5$MeV), we achieve a practical method to evaluate the presence and the amount of these light isotopes in technological materials.
Until now, the only way to do this was through atomic spectroscopy, which, besides not being able to detect the presence of light isotopes, is a much more complex and expensive method.
